ODGI: understanding pangenome graphs

### 2022-03-15 manuscript/main.tex
typos

ODGI improves on prior efforts, based on issues that arose during our work with high-quality \textit{de novo} assemblies that cover almost all parts of the human genome~\citep{Logsdon_2021,Nurk_2021}.
In particular, we find that it is necessary to support graphs with regions of very high numbers of path traversals \REVIEWED{(high depth of path coverage of some nodes)}.
Such motifs can occur in collapsed structures generated by ambiguous sequence homology relationships in repeats found in the centromeres and other segmental duplications.
If we cannot process such regions, \REVIEWED{we cannot understand them}, and \REVIEWED{our only option is to build graphs that do not include them.}
\REVIEWED{Our goal is to build tools that allow for a wide range of uses of pangenome graphs, including cases with potentially high path depth.}
%Neither of these solutions allows us \REVIEWED{to evaluate} their biological \REVIEWED{or technical} features.
%\REVIEWED{This approach  presuppose that we are unable to understand variation in these regions.}
To seamlessly represent such difficult regions, we followed an approach implemented in the dynamic version of the Graph BWT (GBWT)~\citep{Siren:2020} and built a node-centric, dynamic, compressed model of the paths.
This design supports node-local modification and update of the graph, which lets us \REVIEWED{build and modify the graph and its paths} in parallel.

>>>>

ODGI improves on prior efforts, based on issues that arose during our work with high-quality \textit{de novo} assemblies that cover almost all parts of the human genome~\citep{Logsdon_2021,Nurk_2021}.
In particular, we find that it is necessary to support graphs with regions of very high numbers of path traversals \REVIEWED{(high depth of path coverage of some nodes, the so-called node depth)}.
Such motifs can occur in collapsed structures generated by ambiguous sequence homology relationships in repeats found in the centromeres and other segmental duplications.
If we cannot process such regions, \REVIEWED{we cannot understand them}, and \REVIEWED{our only option is to build graphs that do not include them.}
\REVIEWED{Our goal is to build tools that allow for a wide range of uses of pangenome graphs, including cases with potentially high path depth.}
%Neither of these solutions allows us \REVIEWED{to evaluate} their biological \REVIEWED{or technical} features.
%\REVIEWED{This approach  presuppose that we are unable to understand variation in these regions.}
To seamlessly represent such difficult regions, we followed an approach implemented in the dynamic version of the Graph BWT (GBWT)~\citep{Siren:2020} and built a node-centric, dynamic, compressed model of the paths.
This design supports node-local modification and update of the graph, which lets us \REVIEWED{build and modify the graph and its paths} in parallel.
